---
title: "Math_3190_HW2"
author: "Gideon Parry"
date: "2/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

The scan function is uesd to read a text file, with the what = argument being used
to decide how the file is read in to R an example of that is

```{r scan function}
scan("satscores.txt", what = "character")
```
Readlines can be used to well, read lines from a file. It can be used to only read the 
first n lines of a file as well

```{r readlines}
my_txt_ex2 <- readLines(paste("satscores.txt", sep = ""),
                        n = 2)
```
read_html is used to read text from an html page. An example of this is 
```{r read_html}
library(xml2)
Brazil <- read_html("http://en.wikipedia.org/wiki/Brazil_national_football_team")
```

readr is a package that include different ways to read rectangular data, such as:

read_csv(): comma-separated values (CSV) files
read_tsv(): tab-separated values (TSV) files
read_delim(): delimited files (CSV and TSV are important special cases)
read_fwf(): fixed-width files
read_table(): whitespace-separated files
read_log(): web log files

readxl is used to read excel files. 
read_exel(filename, sheet="sheet name") is how to get a specific sheet from the 
excel file
## Question 3



Attaching tidyverse and downloading file

```{r file download}
library(tidyverse)
download.file("http://kenpom.com/cbbga22.txt","cbb.txt")

```

Making it a tibble
```{r }
cbb = read.fwf("cbb.txt", widths = c(10,24,4,22,5,3,20))

cbb = tibble(cbb)
View(cbb)

```

Renaming columns
```{r}
cbb <- rename(cbb, Date = V1)
cbb <- rename(cbb, Away_team = V2)
cbb <- rename(cbb, Away_score = V3)
cbb <- rename(cbb,Home_team = V4)
cbb <- rename(cbb, Home_score = V5)
View(cbb)
```

Adding score difference
```{r}
cbb <- cbb %>%
  mutate(score_diff = Home_score - Away_score)
View(cbb)
```

Trimming and formatting date as date
```{r}
cbb = cbb %>%
  mutate(Date = trimws(Date, which = "both"))

library(lubridate)
cbb <- cbb %>%
  mutate(Date = mdy(Date))
```

Trimming home and away teams
``` {r}

cbb <- cbb %>%
  mutate(Away_team = trimws(Away_team, which = "both"))

cbb <- cbb %>%
  mutate(Home_team = trimws(Home_team, which = "both"))
```

Arranging by location
```{r}
cbb <- cbb %>%
  mutate(V7 = trimws(V7, which = "both"))

cbb <- cbb %>%
  mutate(site = ifelse(is.na(V7),Home_team, V7))

cbb <- arrange(cbb, site)

View(cbb)
```

Removing colums
```{r}
cbb <- cbb %>% select(!(V6 | V7))
```

win percentage for one team
```{r}
suu_wins = cbb %>% summarize(
  win_pct = ((sum(Home_team == "Southern Utah" & score_diff > 0) + sum(Away_team == "Southern Utah" & score_diff < 0))
             /(sum(Home_team == "Southern Utah" | Away_team == "Southern Utah"))
       ), 
      wins = (sum(Home_team == "Southern Utah" & score_diff > 0) + sum(Away_team == "Southern Utah" & score_diff < 0)),
      losses = (sum(Home_team == "Southern Utah" & score_diff < 0) + sum(Away_team == "Southern Utah" & score_diff > 0))
      )   
    
View(suu_wins)
```

function to do it for anyone

```{r}
win_pct <- function(team_name, df){
  team = df %>% summarize(
    team = team_name,
    win_pct = ((sum(Home_team == team_name & score_diff > 0) + sum(Away_team == team_name & score_diff < 0))
               /(sum(Home_team == team_name | Away_team == team_name))),
    wins = (sum(Home_team == team_name & score_diff > 0) + sum(Away_team == team_name & score_diff < 0)),
    losses = (sum(Home_team == team_name & score_diff < 0) + sum(Away_team == team_name & score_diff > 0))
    
     )
  return (team)
}
```

Using that function on SUU

```{r}
suu_wpct <- win_pct("Southern Utah", cbb)
View(suu_wpct)
```

Creating tibble with that info for everyone
```{r}
teams = c()
for (i in 1:length(cbb$Away_team)){
  if (!(cbb$Away_team[i] %in% teams)){
    teams = append(teams, cbb$Away_team[i])
  }
}

teams

all_wpct <- tibble(
  team = character(),
  win_pct = numeric(),
  wins = numeric(),
  losses = numeric(),
)


for(i in 1:length(teams)){
  all_wpct[i,] = win_pct(teams[i],cbb)
}
View(all_wpct)
```
Given team
```{r}
team_games <- function(team_name,df){
  games <- df %>%
    filter(Home_team == team_name | Away_team == team_name)
  games <- games %>%
    mutate(score_diff = ifelse(Away_team == team_name, -score_diff, score_diff))
  games <- games %>%
    mutate(points_for = ifelse(Away_team == team_name, Away_score, Home_score))
  games <- games %>%
    mutate(points_against = ifelse(Away_team == team_name, Home_score, Away_score))
  games <- arrange(games, Date)
  return(games)
}
```

checking that with SUU

```{r}
suu_games <- team_games("Southern Utah", cbb)
View(suu_games)
```

```{r}
library(reshape2)
suu_games <- suu_games %>%
  select(Date | points_for | points_against | score_diff)

View(suu_games)
suu_games <- melt(suu_games, id = "Date")

cols <- c("dark green", "red", "blue")

ggplot(suu_games, aes(x = Date, y = value, color = variable)) +
  geom_line() +
  scale_color_manual(values = cols)

```
Graphing function
```{r}
points_graph <- function(team_name, df){
  games <- df %>%
    filter(Home_team == team_name | Away_team == team_name)
  games <- games %>%
    mutate(score_diff = ifelse(Away_team == team_name, -score_diff, score_diff))
  games <- games %>%
    mutate(points_for = ifelse(Away_team == team_name, Away_score, Home_score))
  games <- games %>%
    mutate(points_against = ifelse(Away_team == team_name, Home_score, Away_score))
  games <- arrange(games, Date)
    
  games <- games %>%
  select(Date | points_for | points_against | score_diff)

  
  games <- melt(games, id = "Date")

  cols <- c("dark green", "red", "blue")
  title <- paste(team_name,"points over time")
  
  return(
  ggplot(games, aes(x = Date, y = value, color = variable)) +
    geom_line() +
    scale_color_manual(values = cols)+
    geom_point()
  + labs(title = title)
  )
}
```


Date filter

```{r}
cbb22 <- cbb %>%
  filter(Date >= as.Date("2022-01-01"))

View(cbb22)
```


Shiny app
```{r}
library(shiny)

teams = c()
for (i in 1:length(cbb$Away_team)){
  if (!(cbb$Away_team[i] %in% teams)){
    teams = append(teams, cbb$Away_team[i])
  }
}

ui <- fluidPage(
  selectInput('xcol', 'Team', teams), 
  plotOutput("graph")
  )

server <- function(input, output){
  output$graph <- renderPlot({
    points_graph(input$xcol,cbb)
  })
  }

shinyApp(server = server, ui = ui)
```
